A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-23 paper.Rmd
done with 4.2.2

As mentioned in Section \ref{power-of-the-tests}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups. Three replications are made for $3 \times 4 = 12$ combinations of common factors $n$ and fitted value distribution, results in $12 \times 3 = 36$ lineups included in data collection period III. In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}. This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated directly from the null model. However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under $H_0$.

@@

All lineups consist of only null plots are planned to be evaluated by 20 participants. 
However, presenting only these lineups to participants are considered to be bad practices as participants will lose interest quickly. 
Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants recruited for data collection period III.

>>>>

Data collection period III was primarily to obtain Rorschach lineup evaluations to estimate $\alpha$ from Equation \ref{eq:pvalue-beta-binomial}, and also to obtain additional evaluations of lineups made with uniform fitted value distribution to ensure consistency in the results. To construct a Rorschach lineup, the data is generated from a model with zero effect size, while the null data are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}. This procedure differs from that of the canonical Rorschach lineup, where all 20 plots are generated directly from the null model, so the method suggested in @vanderplas2021statistical for typical lineups containing a data plot is used to estimate $\alpha$. (The Appendix contains a sensitivity analysis on the effect of uncertainty in the estimation of $\alpha$.) The $3 \times 4 = 12$ treatment levels of the common factors, replicated three times results in 36 lineups. And 6 more evaluations on the 279 lineups with uniform fitted value distribution, results in at least $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants needed.

<!--As mentioned in Section \ref{power-of-the-tests}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null lineups. Three replications are made for $3 \times 4 = 12$ combinations of common factors $n$ and fitted value distribution, results in $12 \times 3 = 36$ lineups included in data collection period III. In these lineups, the data of the data plot is generated from a model with zero effect size, while the data of the 19 null plots are generated using the same simulation method discussed in Section \ref{visual-test-procedure-based-on-lineups}. This generation procedure differs from the canonical Rorschach lineup procedure, which requires that all 20 plots are generated directly from the null model. However, these lineups serve the same fundamental purpose: to assess the number of visually interesting plots generated under $H_0$.

@@

All lineups consist of only null plots are planned to be evaluated by 20 participants. 
However, presenting only these lineups to participants are considered to be bad practices as participants will lose interest quickly. 
Therefore, we plan to collect 6 more evaluations on the 279 lineups with uniform fitted value distribution, result in $(36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133$ participants recruited for data collection period III.
-->

### 2023-04-23 paper.tex
done with 4.2.2

As mentioned in Section \ref{power-of-the-tests}, \(\alpha\) used in
Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using null
lineups. Three replications are made for \(3 \times 4 = 12\)
combinations of common factors \(n\) and fitted value distribution,
results in \(12 \times 3 = 36\) lineups included in data collection
period III. In these lineups, the data of the data plot is generated
from a model with zero effect size, while the data of the 19 null plots
are generated using the same simulation method discussed in Section
\ref{visual-test-procedure-based-on-lineups}. This generation procedure
differs from the canonical Rorschach lineup procedure, which requires
that all 20 plots are generated directly from the null model. However,
these lineups serve the same fundamental purpose: to assess the number
of visually interesting plots generated under \(H_0\).

To account for the fact that our simulation method for these lineups is
not the Rorschach procedure, we use the method suggested in
\citet{vanderplas2021statistical} for typical lineups containing a data
plot to estimate \(\alpha\). (We have included a sensitivity analysis in
the Appendix to examine the impact of the variance of the \(\alpha\)
estimate on our findings.)

All lineups consist of only null plots are planned to be evaluated by 20
participants. However, presenting only these lineups to participants are
considered to be bad practices as participants will lose interest
quickly. Therefore, we plan to collect 6 more evaluations on the 279
lineups with uniform fitted value distribution, result in
\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) participants
recruited for data collection period III.

>>>>

Data collection period III was primarily to obtain Rorschach lineup
evaluations to estimate \(\alpha\) from Equation
\ref{eq:pvalue-beta-binomial}, and also to obtain additional evaluations
of lineups made with uniform fitted value distribution to ensure
consistency in the results. To construct a Rorschach lineup, the data is
generated from a model with zero effect size, while the null data are
generated using the same simulation method discussed in Section
\ref{visual-test-procedure-based-on-lineups}. This procedure differs
from that of the canonical Rorschach lineup, where all 20 plots are
generated directly from the null model, so the method suggested in
\citet{vanderplas2021statistical} for typical lineups containing a data
plot is used to estimate \(\alpha\). (The Appendix contains a
sensitivity analysis on the effect of uncertainty in the estimation of
\(\alpha\).) The \(3 \times 4 = 12\) treatment levels of the common
factors, replicated three times results in 36 lineups. And 6 more
evaluations on the 279 lineups with uniform fitted value distribution,
results in at least
\((36 \times 20 + 279 \times 3 \times 6) / (20-2) = 133\) participants
needed.
